A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-03-26 paper-hierarchical.tex
working...

\begin{abstract}
We consider different approaches to simulate a modern X-ray beamline. Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances. Parameters such as flux, beam size at the image and coherence properties are obtained from simple hand-calculations to sophisticated computer simulations using ray tracing and waveoptics techniques. A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues 
related to highly demagnifying sychrotron sources and exploiting the beam coherence. The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm). This document gives results in term of flux, intensity distribution, focal size, and partial coherence properties using different software tools. The different algorithms and methodologies are discussed with emphasis in benefits and limitations of each one. 
\end{abstract}

@@

1) Introduction to nanofocus and coherent applications

2) ID16a

3) methods and contents

@@

@@

\section{Beamline description}

>>>>

\begin{abstract}
We consider different approaches to simulate a modern X-ray beamline. Several methodologies with increasing complexity are applied to discuss the relevant parameters that quantify the beamline performances. Parameters such as flux, beam size at the image and coherence properties are obtained from simple hand-calculations to sophisticated computer simulations using ray tracing and waveoptics techniques. A last generation x-ray nanofocusing beamline for coherent  applications (ID16A at the ESRF) has been chosen  to study in detail the issues 
related to highly demagnifying sychrotron sources and exploiting the beam coherence. The performance of the beamline is studied for two storages rings: the old ESRF (emittance 4000 pm) and the new ESRF-EBS (emittance 150 pm). In addition to traditional results in term of flux, intensity distribution and focal size, an innovative study on the partial coherence properties based on propagation of coherent modes is presented. The different algorithms and methodologies are implemented in the software suite OASYS. They are discussed with emphasis in benefits and limitations of each one. 
\end{abstract}

@@

Many storage ring based x-ray synchrotron facilities are building or planning upgrades to increase brilliance and coherent flux by one to three orders of magnitude.  The first upgrade of a large facility will be the EBS (Extremely Brilliant Source) \cite{orangebook} at the European Synchrotron Radiation Facility (ESRF), aiming to build a storage ring of 150 pm emittance to significantly boost the associated x-ray coherence. Applications exploiting beam coherence such as such as x-ray photon correlation spectroscopy, coherent diffraction imaging, propagation-based phase contrast imaging and ptychography will strongly benefit from this update.

Accurate calculation and quantitative evaluation of the parameters related to x-ray optics is of paramount importance for designing, building and exploiting the new beamlines. Every modern beamline follows a procedure of conceptual design and then technical design where detailed simulations are essential. The calculations allow the designer to test the design parameters in a virtual computer environment where pros and cons can be studied quantitatively. The selection to the best performant optics configuration requires the detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optical configuration. Deformation in the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformation associated to mechanical issues usually dome using finite eelement analysis. Traditionally the optics calculations for synchrotron beamlines are done using ray tracing techniques. They model the x-ray beam as a collection of different rays that are incoherent one to another. This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc. However, the high coherence of the new sources make necessary to complement the ray tracing methods with models including effects of diffraction and scattering from a coherent beam. A fully coherent beam is still a dream for the new generation storage rings. In spite of a high increate of the coherent fraction of the new beams, the coherent fraction at tipically hard x-ray photon energies is still at few percent for the new sources. This makes still important a ray tracing analysis that does not become obsolete, but has to be complemented with an analysis based on wave optics. Because of the non perfect coherence the wave optics analysis becomes non trivial and can be done with different levels of approximation, as discussed in this work. In addition, wave optics simulations are usually expensive from the computer point of view.

For illustrating the applicability of a battery of methods and solutions for beamline optics simulations we have chosen the ESRF beamline ID16A. This beamline provides a high-brilliance beam focused to a few nanometer. It shows characteristics increasingly searched in modern beamlines: extremely high demagnification (sub-micrometer beam) and high coherence. It combines coherent imaging techniques and X-ray fluorescence microscopy to perform quantitative 3D characterization of the morphology and the elemental composition of specimens in their native state \todo{complete ID16A description}.

@@

The paper is structures as follows: we first introduce in Section~\ref{Bealine description} the main parameters of the beamline and make the selection of the particular choice of the parameters used in the calculations. Section~\label{Simulations} is the central part of the paper, and decribe the different methods used to analyse the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline. In subsection \ref{hand} we calculate with simple methods approximated values of focal size, flux and coherent fraction. Then accuracy is improved using geometrical methods in subsection \ref{ray tracing} that give reasonable results as far as some corrections are considered because of the  mentioned characteristics of high demagnification and high coherence of the studied beamline. The wave optics...

@@

\section{Beamline description}
\label{Beamline description}
